---
title: "SWC R lesson day 1"
author: "Darya Vanichkina"
date: "01/03/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R can be used as a calculator

```{r PEMDAS}
2 + 2
(3+4) * 2
3 + 4 * 2 # comment at end of line
# Comments
2/1000000000000000000000000000
```

## Mathematical functions


```{r MathFunct}
sin(1)
log(1) # natural
log10(10) #base 10
log2(2) # base 2

exp(0.5) # e^1/2
```

## Comparisons
```{r Comparisons}
# equality
1 == 1
# inequality
1 != 1
1 != 2
# less than
1 < 5
# greater than equal to
1 >= 0
# gotcha
2/3
2/3 == 0.66666666666667
all.equal(2/3, 0.66666666666667)
```
# Variables
Names can be 
- camelCase
- with_underscores
- periods.between.words

Names can't
- start with a number (2bad)
- contain spaces (so sorry)
```{r Variables}
x = 10
x <- 10 # this assignment operator
x
x <- 20
x
x <- x + 1
x
```
# Vectorisation
```{r Vectorisation}
1:20
x <- 1:20
length(x)
2 ^ x
```

## Managing your environment
```{r Environment}
# what objects are in my environment
ls()
# remove objects
rm(x)
# where are we in the OS
getwd()
# go somewhere else
setwd(".")
# everything about our environemt
sessionInfo()

```

## Packages
- available on CRAN
- to install packages
```{r Install}
install.packages("packagename")
update.packages()
remove.packages()

library("packagename")
_age
```
- escape kills

```{r Ex2}
mass <- 47.5
age <- 122
mass <- mass * 2.3
age <- age - 20
mass > age
mass
age
library(tidyverse)
library(gapminder)
```
# Getting help
- ?function
- google vignette + package name + function
- StackOverflow
- QuickR
- Rstudio cheat sheets

# Exe

```{r Prim}
c(1, 2, 3)
c('d', 'e', 'f')
c(1, 2, 'f')
```

# R data types

- reading in a csv file

```{r ReadCsv}
cats <- read.csv("cats.csv")
cats$weight
cats$coat
cats$likes_string

paste("My cat is", cats$coat)
```

```{r Types}
typeof(cats$weight)
typeof(cats$coat)
typeof(cats$likes_string)

```

# MAIN TYPES OF DATA

1. double
2. integer
3. complex
4. logical
5. character

```{r Typeof}
typeof(3.14)
typeof(3)
typeof(3L)
typeof(1+1i)
typeof(TRUE)
typeof(F)
typeof("text")
```


```{r Class}
class(cats$coat)
class(cats$weight)
class(cats$likes_string)
class(cats)
```

```{r MakeVector}
myvector <- vector(length = 3)
(my2vector <- vector(mode="integer", length = 3))

as.character(my2vector)
as.numeric(my2vector)
as.logical(my2vector)
str(myvector)
str(cats)
```

```{r COmbineVectors}
combine_vector <- c(2, 6, "c")
str(combine_vector)
as.numeric(combine_vector)
```

```{r WorkingWithData}
(mySeries <- 1:10)
head(mySeries)
tail(mySeries)
str(mySeries)
length(mySeries)
class(mySeries)
typeof(mySeries)x
names(mySeries) <- c(rep("Name1", 5), rep("Name2", 5))
mySeries
```


```{r}
x <- 1:26 
x <- x * 2
names(x) <- LETTERS
names(x) <- letters
```

```{r}
cats2 <- read.csv("cats.csv", stringsAsFactors = FALSE)
```

```{r}
myfactor <- c("Case", "Case", "Control", "Control")
myfactor_asordered <- factor(myfactor, levels = c("Control", "Case"))
myfactor_default <- factor(myfactor)
```


# Session 2

```{r ReadInData}
gapminder <- read.csv("gapminder-FiveYearData.csv")

str(gapminder)
head(gapminder)
typeof(gapminder)
typeof(gapminder$country)
colnames(gapminder)

table(gapminder$country)
table(gapminder$year)
summary(gapminder)
```


## Creating publication quality graphics - Hadley Wickham

```{r Plot}
library(ggplot2)
ggplot(data = gapminder, aes( x = gdpPercap, y = lifeExp)) + geom_point() + theme_minimal()

```
Another library called ggthemes for pretty themes

How has the life expectancy changed over time?

```{r TimeChange}
ggplot(data = gapminder, aes( x = year, y = lifeExp)) + geom_point() + theme_minimal()
# now colour points based on continent column
ggplot(data = gapminder, aes(x = year, y = lifeExp, colour = continent)) + geom_point() + theme_minimal()

# use a line plot instead
ggplot(data = gapminder, aes(x = year, y = lifeExp, colour = continent)) + geom_line() + theme_minimal()

# have both
ggplot(data = gapminder, aes(x = year, y = lifeExp, colour = continent)) + geom_line() + geom_point() + theme_minimal()

```

Overlay stats
```{r}

ggplot(data = gapminder, aes( x = gdpPercap, y = lifeExp)) + geom_point() + theme_minimal()


ggplot(data = gapminder, aes( x = gdpPercap, y = lifeExp)) + geom_point() + theme_minimal() + scale_x_log10()

# add linear model
ggplot(data = gapminder, aes( x = gdpPercap, y = lifeExp)) + geom_point() + theme_minimal() + scale_x_log10() + geom_smooth(method = "lm", size = 1.5)

```
Modify the point colour and the line colour without using aes...

```{r}

ggplot(data = gapminder, aes( x = gdpPercap, y = lifeExp)) + geom_point(colour = "green", size = 2) + theme_minimal() + scale_x_log10() + geom_smooth(method = "lm", size = 1.5, colour= "orange")
```

Practive with the Iris dataset

Test relationship between Sepal.Length and Species 
```{r iris}
iris
# do anova like prism
kruskal.test(Sepal.Length ~ Species, data = iris) 
# 
library(dunn.test)

dunn.test(iris$Sepal.Length, g = iris$Species, method = "bonferroni")

set.seed(1) # set R's random number generator to a specific point
# make a boxplot
#pdf("mytestpdf.pdf")
ggplot(data = iris, aes(x = Species, y = Sepal.Length)) + geom_boxplot() +theme_minimal()+ geom_jitter(aes(colour = Species), width = 0.3) + annotate(geom = "text", x = 1.5, y = 8.25, label = "****" )  + annotate(geom = "text", x = 2.5, y = 8.25, label = "***" )  + annotate("segment", x = 1, xend = 1.9, y = 8.15, yend = 8.15) + annotate("segment", x = 2.1, xend = 3, y = 8.15, yend = 8.15) + annotate("segment", x = 1, xend = 1, y = 8, yend = 8.15)+ annotate("segment", x =1.9, xend = 1.9, y = 8, yend = 8.15)+ annotate("segment", x = 2.1, xend = 2.1, y = 8, yend = 8.15)+ annotate("segment", x = 3, xend = 3, y = 8, yend = 8.15)
#dev.off()
```

```{r SplitPlt}
summary(iris$Petal.Length)
iris$Petal.class <- cut(iris$Petal.Length, c(-Inf, 1.6, 4.35, 5.1, Inf))

ggplot(data = iris, aes(x = Sepal.Width, y = Sepal.Length, colour = Petal.class)) + geom_point() + theme_bw() +facet_grid(Petal.class~ Species)
```


## Back to data frames
```{r AddColumnsAndRows}
cats <- read.csv("cats.csv")
str(cats)

# add a column 
age <- c(1,2,3)
cats <- cbind(cats, age)
str(cats)

newcat <- c("tortoiseshell", 2, 0, 1 )
levels(cats$coat) <- c(levels(cats$coat), "tortoiseshell")
cats <- rbind(cats, newcat)

cats <- rbind(cats, cats)

```

### Subsetting data frames

```{r}
# different ways of parsing
cats <- read.csv("cats.csv")
cats$coat
cats[1]
cats[1:2]
cats[1,1]
cats[1,2]
cats[cats$weight <= 3.2,]
subset.data.frame(cats, cats$weight <= 3.2)
```


```{r Gapminder}
head(gapminder)
head(gapminder[,1:4])
# 
head(gapminder[gapminder$year == 1957,])
# Extract all columns except 1 through to 4
head(gapminder[,-1:-4])
# Extract the rows where the life expectancy is longer the 80years
gapminder[gapminder$lifeExp > 80,]
# Extract the first row, and the fourth and fifth columns (lifeExpand gdpPercap).
gapminder[1, 4:5]
# Advanced: extract rows that contain information for the years2002 and 2007
head(gapminder[gapminder$year == 2002 | 2007,])
head(gapminder[gapminder$year %in% c(2002, 2007),])

head(subset.data.frame(gapminder, gapminder$year %in% c(2002, 2007)))


# europe
subset.data.frame(gapminder, gapminder$continent == "Europe" & gapminder$lifeExp >= 73)

```
# Control flow

```{r}
# foe loop
for(i in 1:10){print(i)}

# if clause
for(x in 1:20){
if (x >= 10){
  print(x)
  print("greater")
}
else if (x > 5)
{  print(x)
  print("medium")}
else {
  print(x)
  print("small")
}
}
```


# Day 4 
## R functions
```{r}
fahr_to_kelvin <- function(temp = 100){
  kelvin <- ((temp - 32) * 5/9) + 273.15
  return(kelvin)
}

# freezing
fahr_to_kelvin(32)
# boiling
fahr_to_kelvin(212)
# 
fahr_to_kelvin()

# write kelvin to celsius function (celsius = kelvin - 273.15)
kelvin_to_celsius <- function(temp){
  celsius <- temp - 273.15
  return(celsius)
}
kelvin_to_celsius(373.15)

# now make another function that combines kelvin_to_celsius AND fahr_to_kelvin to convert from Fahrenheit to Celsius directly
fahr_to_celsius <- function(temp){
  # fahr to kelvin
  kelvin <- fahr_to_kelvin(temp)
  # kelvin to celsius 
  celsius <- kelvin_to_celsius(kelvin)
  # return
  return(celsius)
}

fahr_to_celsius(32)
fahr_to_celsius(212)

fahr_to_celsius2 <- function(temp){
  return(kelvin_to_celsius(fahr_to_kelvin(temp)))
}

fahr_to_celsius2(32)
fahr_to_celsius2(212)
```


```{r GapminderFunc}

calcGDP <- function(dat, year = NULL, country = NULL){
  if(!is.null(year)){
    dat <- dat[dat$year %in% year,]
  }
  if(!is.null(country)){
    dat <- dat[dat$country %in% country, ]
  }
  
  gdp <- dat$pop * dat$gdpPercap
  #print(class(gdp))
  new <- cbind(dat, gdp = gdp)
  
  return(new)
}

head(gapminder)

head(calcGDP(gapminder, year = 2007))
head(calcGDP(gapminder, year = 2007, country = "Afghanistan"))
head(calcGDP(gapminder, year = c(2007,1952), country = "Afghanistan"))

australia <- calcGDP(gapminder, country = "Australia")

write.csv(australia, file = "australia.csv")
# openxlsx
```


Added a line. 







